---
title: "absenteeismR"
output: html_document
---

Had a problem with reshape and tidyverse.

```{r}
install.packages("reshape")
install.packages("reshape2")
install.packages("tidyverse")
```

Load applicable libraries.

```{r}
library(tidyverse)
library(ggplot2)
library(dbplyr)
library(cluster)
library(fpc)
```

Head to check the data table, then remove the "reasonForAbsence.Text" by assigning NULL.
```{r}
head(df)
df$reasonForAbsence.Text. <- NULL
```

Aggregate all data points on primary key, take the mean of all data points for aggregation.
```{r}
df <- aggregate(df, by=list(df$ID), FUN=mean)
head(df)
```

Cluster the project using kmeans, centers determined to be 5 using Elbow Technique
```{r}
cluster <- kmeans(df, centers=5)
plotcluster(df, cluster$cluster)
```

Using qplot for graphing purposes (do not use discrete variables such as Day or Season as they have been averaged due to aggregation.)
```{r}
qplot(x=df$Day.of.the.week, 
      y=df$Distance.from.Residence.to.Work, 
      data=df, 
      colour = df$Day.of.the.week, 
      geom = c("point", "smooth"),
      main = "Absenteeism by Day",
      xlab = "Day", 
      ylab = "Absenteeism in hours")
```

Elbow method for deciding clustering volume, 4 through 6 is applicable.
```{r}
#Elbow Method for finding the optimal number of clusters
set.seed(123)
# Compute and plot wss for k = 2 to k = 15.
k.max <- 15
wss <- sapply(1:k.max, function(k){kmeans(df, k, nstart=50,iter.max = 15 )$tot.withinss})
wss
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```

Head function to do root-cause on reasons for each of the clusters.
```{r}
head(cluster)
```

Adding the clustering groups back into the dataset
```{r}
kMeansResult <- factor(cluster$cluster)
df2 <- df
df2$cluster <- kMeansResult
head(df2)
```

